### 2022-11-22 docs/paper.tex
add Fréchet distance to intro section

\documentclass[
oneside,
fontsize=11pt
]{scrartcl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% packages
%%%

%%%
%%% encoding and language set
%%%

\usepackage[english]{babel}
\usepackage[a4paper]{geometry}

%%% fontenc, ae, aecompl: coding of characters in PDF documents
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage[autostyle=true,german=quotes]{csquotes}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{minted}

%%%
%%% technical packages
%%%

%%% amsmath, amssymb, amstext: support for mathematics
\usepackage{amsmath,amssymb,amstext}

%%%
%%% Colors
%%%
\RequirePackage[table,dvipsnames]{xcolor}
\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}

%%%
%%% Macros
%%%
\newtheorem{mydef}{Definition}

%%%
%%% PDF
%%%

\usepackage{ifpdf}

%%% Should be LAST usepackage-call!
%%% For docu on that, see reference on package ``hyperref''
\ifpdf%   (definitions for using pdflatex instead of latex)

  %%% graphicx: support for graphics
  \usepackage[pdftex]{graphicx}

  \pdfcompresslevel=9

  %%% hyperref (hyperlinks in PDF): for more options or more detailed
  %%%          explanations, see the documentation of the hyperref-package
  \usepackage[%
    %%% general options
    pdftex=true,      %% sets up hyperref for use with the pdftex program
    %plainpages=false, %% set it to false, if pdflatex complains: ``destination with same identifier already exists''
    %
    %%% extension options
    backref,      %% adds a backlink text to the end of each item in the bibliography
    pagebackref=false, %% if true, creates backward references as a list of page numbers in the bibliography
    colorlinks=true,   %% turn on colored links (true is better for on-screen reading, false is better for printout versions)
    linkcolor=TUGreen,
    urlcolor=TUGreen,
    %
    %%% PDF-specific display options
    bookmarks=true,          %% if true, generate PDF bookmarks (requires two passes of pdflatex)
    bookmarksopen=true,     %% if true, show all PDF bookmarks expanded
    bookmarksnumbered=true, %% if true, add the section numbers to the bookmarks
    %pdfstartpage={1},        %% determines, on which page the PDF file is opened
    pdfpagemode=None         %% None, UseOutlines (=show bookmarks), UseThumbs (show thumbnails), FullScreen
  ]{hyperref}

  %%% provide all graphics (also) in this format, so you don't have
  %%% to add the file extensions to the \includegraphics-command
  %%% and/or you don't have to distinguish between generating
  %%% dvi/ps (through latex) and pdf (through pdflatex)
  \DeclareGraphicsExtensions{.pdf}

\else %else   (definitions for using latex instead of pdflatex)

  \usepackage[dvips]{graphicx}

  \DeclareGraphicsExtensions{.eps}

  \usepackage[%
    dvips,           %% sets up hyperref for use with the dvips driver
    colorlinks=false %% better for printout version; almost every hyperref-extension is eliminated by using dvips
  ]{hyperref}

\fi

%%% sets the PDF-Information options
%%% (see fields in Acrobat Reader: ``File -> Document properties -> Summary'')
%%% Note: this method is better than as options of the hyperref-package (options are expanded correctly)
\hypersetup{
  pdftitle={Fréchet Distance}, %%
  pdfauthor={Tom Stein}, %%
  pdfsubject={Seminar Algorithm Engineering 22/23}, %%
  pdfcreator={Accomplished with LaTeX2e and pdfLaTeX with hyperref-package.}, %% 
  pdfproducer={}, %%
  pdfkeywords={} %%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% define the titlepage
%%%

% \subject{}   %% subject which appears above titlehead
% \titlehead{} %% special heading for the titlepage

%%% title
\title{Fréchet Distance}

%%% author(s)
\author{Tom Stein}

%%% date
\date{November 27, 2022}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% begin document
%%%

\begin{document}

% \pagenumbering{roman} %% small roman page numbers

%%% include the title
% \thispagestyle{empty}  %% no header/footer (only) on this page
%  \maketitle

% Titlepage ---------------------------------------------------------
%
\pdfbookmark{Titelpage}{pdf:title}
\include{titlepage}

%%% start a new page and display the table of contents
% \newpage
% \tableofcontents

%%% start a new page and display the list of figures
% \newpage
% \listoffigures

%%% start a new page and display the list of tables
% \newpage
% \listoftables

%%% display the main document on a new page 
\newpage

% \pagenumbering{arabic} %% normal page numbers (include it, if roman was used above)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% begin main document
%%% structure: \section \subsection \subsubsection \paragraph \subparagraph
%%%

\newgeometry{left=3cm, right=4.5cm, top=4cm, bottom=4cm}

\section*{Abstract}

\section{Introduction}
% motivation
At first it seems that measuring the similairty between two curves
is a very theoretical problem
but it has many, not that obvious, pratical applications.
It is used in Geo-Information-Systems (GIS) to 
quantify the similarity or quality of two maps, % cite
find patterns in vehicle traffic or animal tracking data 
by trajectory analysis, % cite
gesture recognition on touch devices like smartphones \cite{hu_research_2022}, % cite
and shape matching in the field of computer vision.

To quantify the similairty of two given curves (e.g. shapes, paths, or trajectories)
a distance metric can be used. 
We can't simply use a distance metric like eucledian distance 
becuase we don't want to compare points to points 
but instead compare curves to curves.

% two (polygonal) curves
% TODO: replace continous funcitons with other name
\begin{mydef}
  \label{def_curve}
  We define a curve as a continuous function $f : [a,b] \mapsto V$ 
  with $a,b \in \mathbb{R}, a \leq b$ and $V$ being the space of the curve,
  e.g., $\mathbb{R}^2$ or $\mathbb{R}^3$.
\end{mydef}

\begin{mydef}
  \label{def_polygonal_curve}
  We define a polygonal curve (polygonal chain) 
  as a linearly interpolated curve 
  on $n \in \mathbb{N}$ points in $V$ such that
  $P : [0,n] \mapsto V$ 
  with  $P(i + \lambda) = (1 - \lambda)P(i) + \lambda P(i+1)$
  for all $\lambda \in [0,1]$ and all $i \in \{0, 1, \dots, n-1\}$.
\end{mydef}

In the following we will focus on polygonal curves (polygonal chains) 
instead of classical continuous curves, ofen used in pure mathematics, 
because most data we have in this area is 
captured by some none continuous mechanism, e.g., GPS positions from a tracking device.
Addionally, we focus on the 2D Space, i.e. $V = \mathbb{R}^2$.
However, many concepts also apply to normal curves and higher dimensional spaces.

% Hausdorff distance
The commonly known \textit{Hausdorff distance} can be understood 
as the largest distance between any two points on the curves. 
For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}. \cite{alt_computing_1995}

\begin{mydef}
  \label{def_hausdorff_distance}
  Given two (polygonal) curves $P$ and $Q$, 
  their Hausdorff distance $\delta_{hd}(P,Q)$ is mathematically defined as: 
  $$\delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)$$
\end{mydef}

This metric is very intuitive and can be easily computed for polygonal curves \cite{ko_complexity_2013}.
It finds applications in computer vision and computer graphics 
where it is used to compare the similarity of two mesh objects \cite{cignoni_metro_1998,ko_complexity_2013}. 
However, this metric does not take the course of the curves into account, 
but only the point sets of the individual curves.
Two curves may have a small Hausdorff-distance but seem very dissimilar 
when looking at them, see \autoref{fig_hausdorff_distance_bad_example}.

\begin{figure}[ht]
  % TODO: insert figure
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
      }
      \caption{Hausdorff distance between the red curve $P$ and green curve $Q$. }
      \label{fig_hausdorff_distance_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
      }
      \caption[Hausdorff distance of two dissimilar curves]{Two dissimilar curves with a small Hausdorff distance.}
      \label{fig_hausdorff_distance_bad_example}
  \end{subfigure}
  \caption[Hausdorff distance examples]{Hausdorff distance examples.}
  \label{fig_hausdorff_distance_examples}
\end{figure}

% Fréchet distance motivation
In reality, the course of the curves is important for many applications. 
For example, in handwriting character recognition on digital input devices, especially in Chinese, % TODO: add more citations
gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
For this reason, the \textit{Fréchet distance}, named after the famous french mathematician René Maurice Fréchet, was introduced.

A common analogy for the Fréchet distance can be shown by a man walking his dog. 
Assume that both the man and the dog are moving along a predefined fixed curve. 
Both are free to adjust their speed as they wish, even to stop. 
However, they are not allowed to go backwards. 
We are looking for the minimum length of the dog leash 
that is necessary for both of them to reach their destination. 
The length of the dog leash is then exactly the Fréchet distance of the two curves.

% continuous Fréchet distance equation

\begin{mydef}
  \label{def_frechet_distance}
  Given two (polygonal) curves $P: [a,a'] \mapsto V$, $Q: [b,b'] \mapsto V$ and a distance function $\text{dist}: (V,V) \mapsto \mathbb{R}$, 
  their Fréchet distance $\delta_{F}(P,Q)$ is mathematically defined as: 
  $$\delta_{F}(P,Q) = \inf_{\substack{\alpha: [0,1] \mapsto [a, a'] \\ \beta: [0,1] \mapsto [b, b']}} \max_{t \in [0,1]} \text{dist}(P(\alpha(t)), Q(\beta(t)))$$
  where $\alpha$ and $\beta$ represend the walking speed as 
  continuous monotonically increasing functions with
  $\alpha(0) = a, \alpha(1) = a'$, 
  $\beta(0) =b, \beta(1) =b'$.
\end{mydef}

% TODO: Add example of Frechet distance

% Scope
% - Different Variants and their differences
% - Original Algorithm by Alt and Godau
% - Original Algorihm for discrete Fréchet distance
% - Newer Algorithms (vor variants)

% Fréchet distance Variants 
\section{Variants}
Through the years many slightly differnt variants of the continuous Fréchet distance (see Definition \ref{def_frechet_distance}) have evolved. 
The continuous Fréchet distance takes into account every single point on the curves, 
with interpolation in the case of polygonal curves, 
without allowing to walk backwards 
and without any obstacles in between the two curves.

% TODO: Why allow it? Better algorithms?
The \textit{weak Fréchet distance} slightly modifies the continuous version by allowing 
going backwards.

% TODO: Is that mapping bijective? Add sources
The \textit{discrete Fréchet distance} 
This gives an upper bound on the continuous Fréchet distance % Source?
and is often used as a approximation. % Source?
% discrete Fréchet distance and contrast to hausdorff distance

% (Geodesic Fréchet distance)

% (Homotopic Fréchet distance)

\section{Algorithms}
% Chapter summary and structure

\subsection{Original algorithm for the continuous Fréchet distance}
% Free Space Diagram
% Runtime analysis (?)
% Correctness proof (?)

\subsection{Improved Version?}

\subsection{Algorithms for different variants}
% discrete Fréchet distance

% weak Fréchet distance

\section{Conclusion}

%%%
%%% end main document
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\appendix  %% include it, if something (bibliography, index, ...) follows below

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% bibliography
%%%
%%% available styles: abbrv, acm, alpha, apalike, ieeetr, plain, siam, unsrt
%%%
\bibliographystyle{alpha}

%%% name of the bibliography file without .bib
%%% e.g.: literature.bib -> \bibliography{literature}
\bibliography{literature}

\end{document}
%%% }}}
%%% END OF FILE